---
title: "BSPv2"
author: "Aleksander Nitka"
date: "25/09/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(R.matlab)
library(ggplot2)
library(knitr)
library(ggpubr)
```

## List matlab files in directory
```{r, list matlab files in directory}

matdir = "/Users/aleksandernitka/OneDrive\ -\ The\ University\ of\ Nottingham/Lehigh_2018/Data/matlab_data"

files.mat = list.files(path = matdir, pattern = "\\.mat$")
print(files.mat)
```

## Load files into a one data frame
```{r, load files}
for (i in 1:length(files.mat)) {
    
    # Extract details from file name
    exp = strsplit(files.mat[i], "_")[[1]][1]
    ss = strsplit(files.mat[i], "_")[[1]][2]
    block = strsplit(files.mat[i], "_")[[1]][3]
    cond = strsplit(strsplit(files.mat[i], "_")[[1]][4], '[.]')[[1]][1]

    # Read mat file
    dat = readMat( paste(matdir, files.mat[i], sep = "/"))
    
    # Extract data from mat to R
    # searchRT - rection time to non-porbe trials, 0 if probe trial but 0 also if no response
    # targetletters - no of accurately reported letters on target side
    # distletters - no of accurately reported letters on non-target side
    # probe - logical if trial was a probe or not
    # SearchRespCode - 38 is UpArrow and 40 is DownArrow
    
    # TODO
    # How do I know if response to rotation was correct?
    
    tmp = as.data.frame(cbind(t(dat$searchRT), 
                              t(dat$searchRespCode),
                              t(dat$targletters),
                              t(dat$distletters),
                              dat$probe, 
                              t(dat$FA), 
                              t(dat$colorcue),
                              exp, ss, block, cond
                              ))
    
    names(tmp) = c('RT', 'SearchRespCode', 'TargLetters', 'DistLetters', 'ProbeTrial', 'FA', 
                   'ColorCue', 'Exp', 'SS', 'Block', 'Cond')
    
    # Extract orientation codes
    orient = as.data.frame(dat$Orient)
    names(orient) = c('orient01', 'orient02', 'orient03', 'orient04',
                      'orient05', 'orient06', 'orient07', 'orient08',
                      'orient09', 'orient10', 'orient11', 'orient12')
    
    tmp = cbind(tmp, orient)
    
    if (i == 1){
        # Create main data frame by copying the tmp
        BSPv2 = tmp
        print('BSPv2 data frame created')
        print(sprintf('Added ss %s, %s, %s to BSPv2', ss, block, cond))
        remove(tmp, block, cond, exp, ss, orient, dat)
    }
    else {
        # bind main data frame and tmp 
        BSPv2 = rbind(BSPv2, tmp)
        print(sprintf('Added ss %s, %s, %s to BSPv2', ss, block, cond))
        remove(tmp, block, cond, exp, ss, orient, dat)
        
    }
    
}
```

## Do some data manipulations, prep for mean analysis
```{r, prepare df for analysis}
# data type must be changed from integer to numeric, but converting directly changes the values,
# must convert to string and then to numeric to keep the actual values
BSPv2$RT = as.numeric(as.character(BSPv2$RT))
BSPv2$TargLetters = as.numeric(BSPv2$TargLetters)
BSPv2$DistLetters = as.numeric(BSPv2$DistLetters)

# If RT = 0 and trial != Probe then it was a timeOut
# need to mark those as otherwise 0 will be counted towards the mean
BSPv2$timeOut = 0
BSPv2$timeOut[BSPv2$RT == 0 & BSPv2$ProbeTrial == 0] = 1

# Get subject IDs
SSid = as.numeric(unique(BSPv2$SS))

# Create a DF for means
means_BSPv2 = data.frame(matrix(nrow = length(SSid)*3, ncol = 5))
names(means_BSPv2) = c('SS', 'Cond', 'mRT', 'LettersTarget', 'LettersDistractor')
means_BSPv2$SS = SSid
means_BSPv2$Cond = c(rep('cond0',length(SSid)), rep('cond1',length(SSid)), rep('cond2',length(SSid)))
```


## Extract means
```{r, extract means}
for (i in 1:nrow(means_BSPv2)){
    
    means_BSPv2$mRT[i] = mean(subset(BSPv2$RT, 
                                     BSPv2$SS == means_BSPv2$SS[i] & 
                                         BSPv2$Cond == means_BSPv2$Cond[i] & 
                                         BSPv2$ProbeTrial == 0 &
                                         BSPv2$timeOut == 0))
    
    means_BSPv2$LettersTarget[i] = mean(subset(BSPv2$TargLetters, 
                                     BSPv2$SS == means_BSPv2$SS[i] & 
                                         BSPv2$Cond == means_BSPv2$Cond[i] & 
                                         BSPv2$ProbeTrial == 1))
    
    means_BSPv2$LettersDistractor[i] = mean(subset(BSPv2$DistLetters, 
                                               BSPv2$SS == means_BSPv2$SS[i] & 
                                                   BSPv2$Cond == means_BSPv2$Cond[i] & 
                                                   BSPv2$ProbeTrial == 1))
    
}

# Make sure that data types as good
means_BSPv2$Cond = as.factor(means_BSPv2$Cond)
means_BSPv2$LettersTarget = as.numeric(means_BSPv2$LettersTarget)
means_BSPv2$LettersDistractor = as.numeric(means_BSPv2$LettersDistractor)


# Probe trials, Create Difference Index = Target - Distractor

means_BSPv2$LettersDiff = means_BSPv2$LettersTarget - means_BSPv2$LettersDistractor

```

## Individual Means
```{r, indi means, echo=0}
kable(means_BSPv2)

```

##Fast results

###RT analysis 
Performed for N = `r length(SSid)`. Mean RT for **'cond0'** was `r mean(subset(means_BSPv2$mRT, means_BSPv2$Cond == 'cond0'))` (SD = `r sd(subset(means_BSPv2$mRT, means_BSPv2$Cond == 'cond0'))`), for **'cond1'** it was `r mean(subset(means_BSPv2$mRT, means_BSPv2$Cond == 'cond1'))` (SD = `r sd(subset(means_BSPv2$mRT, means_BSPv2$Cond == 'cond1'))`) and for **'cond2'** it was `r mean(subset(means_BSPv2$mRT, means_BSPv2$Cond == 'cond2'))` (SD = `r sd(subset(means_BSPv2$mRT, means_BSPv2$Cond == 'cond2'))`).

###Probe analysis. 
On **'cond0'** trials participants reported M = `r mean(means_BSPv2$LettersTarget[means_BSPv2$Cond == 'cond0'])` (SD = `r sd(means_BSPv2$LettersTarget[means_BSPv2$Cond == 'cond0'])`) letters on the Target Side and M = `r mean(means_BSPv2$LettersDistractor[means_BSPv2$Cond == 'cond0'])` (SD = `r sd(means_BSPv2$LettersDistractor[means_BSPv2$Cond == 'cond0'])`) on Distractor Side. During the **'cond1'** trials participants reported M = `r mean(means_BSPv2$LettersTarget[means_BSPv2$Cond == 'cond1'])` (SD = `r sd(means_BSPv2$LettersTarget[means_BSPv2$Cond == 'cond1'])`) letters on the Target Side and M = `r mean(means_BSPv2$LettersDistractor[means_BSPv2$Cond == 'cond1'])` (SD = `r sd(means_BSPv2$LettersDistractor[means_BSPv2$Cond == 'cond1'])`) on Distractor Side. Finally, for the **'cond2'** trials participants reported M = `r mean(means_BSPv2$LettersTarget[means_BSPv2$Cond == 'cond2'])` (SD = `r sd(means_BSPv2$LettersTarget[means_BSPv2$Cond == 'cond2'])`) letters on the Target Side and M = `r mean(means_BSPv2$LettersDistractor[means_BSPv2$Cond == 'cond2'])` (SD = `r sd(means_BSPv2$LettersDistractor[means_BSPv2$Cond == 'cond2'])`) on Distractor Side.

```{r, plots - distributions, echo=0}
# RT distributions for all subjects

for (i in 1:length(SSid)){
    
    if (i == 1){
        plot(density(BSPv2$RT[BSPv2$SS == SSid[i] & BSPv2$Cond == 'cond0' & BSPv2$ProbeTrial == 0 & BSPv2$timeOut == 0]), ylim = c(0,1), col = 'red', main = 'RT distribution, subject/condition')
        lines(density(BSPv2$RT[BSPv2$SS == SSid[i] & BSPv2$Cond == 'cond1' & BSPv2$ProbeTrial == 0 & BSPv2$timeOut == 0]), ylim = c(0,1), col = 'blue')
        lines(density(BSPv2$RT[BSPv2$SS == SSid[i] & BSPv2$Cond == 'cond2' & BSPv2$ProbeTrial == 0 & BSPv2$timeOut == 0]), ylim = c(0,1), col = 'green')
        
        legend(3.5,1, c('cond0','cond1', 'cond2'), lty=c(1,1), lwd=c(2.5,2.5), col=c('red','blue','green'))
    }
    
    else {
        lines(density(BSPv2$RT[BSPv2$SS == SSid[i] & BSPv2$ProbeTrial == 0 & BSPv2$timeOut == 0 ]), col = 'red')
                lines(density(BSPv2$RT[BSPv2$SS == SSid[i] & BSPv2$Cond == 'cond1' & BSPv2$ProbeTrial == 0 & BSPv2$timeOut == 0]), ylim = c(0,1), col = 'blue')
        lines(density(BSPv2$RT[BSPv2$SS == SSid[i] & BSPv2$Cond == 'cond2' & BSPv2$ProbeTrial == 0 & BSPv2$timeOut == 0]), ylim = c(0,1), col = 'green')
    }
    
}

```

```{r, ggplot - means}
#ggplot(data = means_BSPv2, aes(Cond, mRT)) + geom_point(color="skyblue") + stat_summary(fun.y=mean,geom="line",lwd=1,aes(group=1))


p = ggboxplot(means_BSPv2, x = "Cond", y = "mRT",
                color = "Cond", palette =c("dark red", "dark green", "dark blue"),
                add = "jitter", shape = "Cond")

my_test = list( c("cond0", "cond1"), c("cond1", "cond2"), c("cond0", "cond2") )

p + stat_compare_means(comparisons = my_test, method = "t.test") + ggtitle("RT")


ggdensity(means_BSPv2, x = "mRT",
   add = "mean", rug = TRUE,
   color = "Cond", fill = "Cond",
   palette = c("dark red", "dark green", "dark blue"))



```


```{r, Assumptions of ANOVA test}
## RT
# normality test
shapiro.test(means_BSPv2$mRT)
# Bartlett Test of Homogeneity of Variances PARAMETIC
bartlett.test(mRT~Cond, data=means_BSPv2)
# Figner-Killeen Test of Homogeneity of Variances NON-PARAMETIC
fligner.test(mRT~Cond, data=means_BSPv2) 
```